# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

``` {r read, echo=TRUE}
activity<-read.csv("activity.csv", na.strings = "NA", colClasses=c("integer", "Date", "integer"))

activity$datetime<-strptime(paste(activity$date, sprintf("%04d", activity$interval)), "%Y-%m-%d %H%M")
```
head(activity)


## What is mean total number of steps taken per day?
``` {r Stepsperday, echo=TRUE}
Stepsperday <- aggregate(steps~date, data=activity, sum)  
hist(Stepsperday$steps, main="Histogram of steps per day", xlab = "Steps per day")
stepsperday.mean<-mean(Stepsperday$steps)
stepsperday.median<-median(Stepsperday$steps)
```
The mean value of the total number of steps taken per day is `r stepsperday.mean`.

The median value of the total number of steps taken per day is `r stepsperday.median`.


## What is the average daily activity pattern?
``` {r Daily, echo=TRUE}
activity$timeoftheday <- strptime(sprintf("%04d", activity$interval), "%H%M")
library(plyr)
meanbytimeoftheday <- ddply(activity, .(timeoftheday), summarize, mean = mean(steps, na.rm = TRUE))
plot(meanbytimeoftheday$timeoftheday, meanbytimeoftheday$mean, type="l", main="Average number of steps taken per 5-minute interval", xlab="5-minute interval", ylab="Average number of steps")
maxsteps<-round(max(meanbytimeoftheday$mean),2)
maxinterval<-meanbytimeoftheday$timeoftheday[which.max(meanbytimeoftheday$mean)]
maxinterval.nodate<-strftime(maxinterval, "%H:%M")
```

The maximum number of steps on average across all the days in the dataset is `r maxsteps` and it occurs at `r maxinterval.nodate`.
    
## Imputing missing values
``` {r Imputing, echo=TRUE}
na <- is.na(activity$steps) 
sum(na)
filledactivity<-activity
for (i in which(is.na(filledactivity$steps))) {
filledactivity$steps[i]<-meanbytimeoftheday$mean[(meanbytimeoftheday$timeoftheday==filledactivity$timeoftheday[i])]
}
filledstepsperday <- aggregate(steps~date, data=filledactivity, sum)  
hist(filledstepsperday$steps, main="Histogram of steps per day", xlab = "Steps per day")
filledstepsperday.mean<-mean(filledstepsperday$steps)
filledstepsperday.median<-median(filledstepsperday$steps)
```
The mean value of the total number of steps taken per day is `r filledstepsperday.mean`. The mean value is the same as before imputation.

The median value of the total number of steps taken per day is `r filledstepsperday.median`. The median value has changed after imputation.


## Are there differences in activity patterns between weekdays and weekends?
``` {r weekend, echo=TRUE}
filledactivity$weekend<-rep(0, length(filledactivity$date))
filledactivity$weekend[(as.POSIXlt(filledactivity$date)$wday==0 | as.POSIXlt(filledactivity$date)$wday==6)]<-1
weekdayactivity<-subset(filledactivity, weekend==0)
weekendactivity<-subset(filledactivity, weekend==1)
meanweekend <- ddply(weekendactivity, .(interval), summarize, mean = mean(steps, na.rm = TRUE), weekend=1)
meanweekday <- ddply(weekdayactivity, .(interval), summarize, mean = mean(steps, na.rm = TRUE), weekend=0)
mean2<-rbind(meanweekend, meanweekday)
mean2$weekend <- factor(mean2$weekend, labels = c("weekday", "weekend"))
library(lattice)
xyplot(mean ~ interval | weekend, data=mean2, type="l", xlab="Interval", ylab="Number of steps", layout = c(1, 2)) ## Plot with 2 panels
```
